---
title: "RFDS_CH2"
output: html_document
---

```{r libraries}
library(tidyverse)
library(ggplot2)
```


Alt+- is a shortcut for assignment symbol <-. This works as an r function. When I tried typing in this (text) area, it showed as a = sign. Once I tried in the console or in an r chunk, it worked.

Insert an r chunk ctrl+Alt+i

ctrl+(up arrow) lists all commands you've typed that start with those letters.

If you want to print an assignment when you assign it, wrap it in parenthesis (my_variable <- 10)

##### pg. 40

#1

Q: Why does this code not work?

```{r pg40_1}
my_variable <- 10
my_varıable
```

Works in this case? but shouldn't because the i is different. Need to be specific!

#2

Q: Tweak each of the following R commands so that they run correctly:

ggplot(dota = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))

fliter(mpg, cyl = 8)
filter(diamond, carat > 3)

```{r pg40_2}
ggplot(data = mpg) + # dota -> data
  geom_point(mapping = aes(x = displ, y = hwy))
```

```{r pg_40_2b}
filter(mpg, cyl == 8) # fliter -> filter, and '==' instead of '='
```

```{r pg_40_2c}
filter(diamonds, carat > 3) # dataset is name diamonds
```

#3

Q: Press Alt+Shift+K. What happens? How can you get to the same place using the menus?

A: Keyboard Shortcuts Menu
For more shortcuts, Tools > Keyboard Shortcuts Help

