---
title: "RFDS_CH3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(nycflights13)
library(tidyverse)
library(ggplot2)
```

When comparing two vectors for equality, use near(). "This is a safe way of comparing if two vectors of floating point numbers are (pairwise) equal. This is safer than using ==, because it has a built in tolerance"

When using filter(), & can be the same as a ,

```{r example_a}
filter(flights, arr_delay <= 120, dep_delay <= 120)
```
```{r example_a_1}
filter(flights, arr_delay <= 120 & dep_delay <= 120)
```


##### pg. 49

#1

Q: Find all flights that:

a. Had a n arrival delay of two or more hours

```{r pg49_1a}
filter(flights, arr_delay >= 120)
```

b. Flew to Houston (IAH or HOU)

```{r pg49_1b}
filter(flights, dest %in% c('IAH', 'HOU'))
```

c. Were operated by United, American or Delta

```{r pg49_1c}
filter(flights, carrier %in% c('UA', 'AA', 'DL'))
```

d. Departed in summer (July, August, and September)

```{r pg49_1d}
filter(flights, month %in% c('7', '8', '9'))
```

e. Arrived more than two hours late, but didn’t leave late

```{r pg49_1e}
filter(flights, arr_delay > 120 & dep_delay <= 0)
```

f. Were delayed by at least an hour, but made up over 30 minutes in flight

```{r pg49_1f}
filter(flights, dep_delay >= 60 & dep_delay - arr_delay > 30)
```

g. Departed between midnight and 6am (inclusive)

check min and max of dep_time

```{r pg49_1g}
filter(flights, dep_time <= 600 | dep_time == 2400)
```

#2

Q: Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?

A: ?between = "This is a shortcut for x >= left & x <= right, implemented efficiently in C++ for local values, and translated to the appropriate SQL for remote tables.
Usage
between(x, left, right)"
It would work for 1d

#3

Q: How many flights have a missing dep_time? What other variables are missing? What might these rows represent?

```{r pg49_3}
filter(flights, is.na(dep_time))
```

A: dep_delay, arr_time, arr_delay are all missing. They may be cancelled flights.

#4

Q: Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)

A: Follows mathematical rules, so the value does not matter in the examples given. For example - anything raised to 0 is going to be 1.

##### pg. 51

#1

Q: How could you use arrange() to sort all missing values to the start? (Hint: use is.na().)

A: From : https://jrnold.github.io/r4ds-exercise-solutions/transform.html#arrange-rows-with-arrange
  To put NA values first, we can add an indicator of whether the column has a missing value. Then we sort by the missing indicator column and the column of interest. For example, to sort the data frame by departure time (dep_time) in ascending order but NA values first, run the following.
  
```{r pg51_1}

arrange(flights, desc(is.na(dep_time)), dep_time)

```

#2

Q: Sort flights to find the most delayed flights. Find the flights that left earliest.

  
```{r pg51_2a}

# most delayed

arrange(flights, desc(dep_delay))

```

```{r pg51_2b}

# earliest departing

arrange(flights, dep_delay)

```

#3

Q: Sort flights to find the fastest flights.

A: You have to create a new column with the "ground speed". Then sort descending by ground speed.
  
```{r pg51_3}

fastest_flights <- mutate(flights, mph = (distance / air_time) * 60)

head(arrange(fastest_flights, desc(mph)))

```

#4

Q: Which flights traveled the longest? Which traveled the shortest?

  
```{r pg51_4a}

# traveled the longest

arrange(flights, desc(air_time))

```

```{r pg51_4b}

# traveled the shortest

arrange(flights, air_time)

```

##### pg. 54

#1

Q: Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.
  
```{r pg54_1}

select(flights, dep_time, dept_delay, arr_time, arr_delay)

select(flights, 4, 6, 7, 9)

select(flights, one_of(c("dep_time", "dep_delay", "arr_time", "arr_delay")))

select(flights, starts_with("dep_"), starts_with("arr_"))

select(flights, matches("^(dep|arr)_(time|delay)$"))

```

#2

Q: What happens if you include the name of a variable multiple times in a select() call?

A: The select() call ignores duplication. The duplicated variable will appear only once in the results in the first position it is called.

#3

Q: What does the one_of() function do? Why might it be helpful in conjunction with this vector?

vars <- c("year", "month", "day", "dept_delay", "arr_delay")

A: one_of() "Matches variable names in a character vector." (from documentation). This can be helpful because it is easier to create a character vector of variable names.

In newer versions of dplyr, the one_of() is not needed. You have to be cautious of whether the name of the vector containing the variable names is not itself a column name.

#4

Q: Does the result of running the following code surprise you? How do the select helpers deal with case by detault? How can you change the default?

A: It didn't at first, but I realize the case should matter? The default from documentation shows ignore.case	
If TRUE, the default, ignores case when matching names. You could change this to FALSE if needed.

##### pg. 58

#1

Q: Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.
  
```{r pg58_1}

# Solutions has great explanation. Basically breaking up the hours and minutes from the time, converts the hours to minutes, then the %% 1440 accounts for midnight ("2400", but should be 0).
# The %% 1400 works because operation takes the remainder of the number of minutes/1440. If the number of minutes turns out to be 1440 (the number of minutes calculated if time is 2400) then the remainder is zero. If it is anything less than 1440 minutes (every other time), it will return that number (the remainder). Very tricky.

flights_times <- mutate(flights,
  dep_time_mins = (dep_time %/% 100 * 60 + dep_time %% 100) %% 1440,
  sched_dep_time_mins = (sched_dep_time %/% 100 * 60 +
    sched_dep_time %% 100) %% 1440
)

```

#2

Q: Compare air_time with arr_time - dep_time. What do you expect to see? What do you see? What do you need to do to fix it?

A: I do not expect them to be the same because of time zones and flights spanning midnight. Soutions has a great explanation.

```{r pg58_2}

flights_with_arr_minus_dep <- mutate(flights, arr_minus_dep = arr_time - dep_time)

select(flights_with_arr_minus_dep, arr_minus_dep, air_time)

```

#3

Q: Compare dep_time, sched_dep_time, and dep_delay. How would you expect those three numbers to be related?

A: I would expect sched_dep_time - dep_delay to equal dep_time.

#4

Q: Find the 10 most delayed flights using a ranking function. How do you want to handle ties? Carefully read the documentation for min_rank().

A: There are no ties in the top 10, but I would still use min_rank rather than dense_rank because you could more easily see the top 10 results by locating the rank with 10th place (assuming it is not a tie). dense_rank could have ties and would then continue the rank with the next space, so if you grabbed the 1-10 ranked rows, you could have more than 10 flights.

```{r pg58_4}

flights_with_min_rank <- mutate(flights, min_rank_dep_delay = min_rank(desc(dep_delay)))
arrange(flights_with_min_rank, min_rank_dep_delay)

```

#5

Q: What does 1:3 + 1:10 return? Why?

A: Gives a warning: longer object length is not a multiple of shorter object length [1]  2  4  6  5  7  9  8 10 12 11
The shorter vector is being recycled to complete the addition with the longer vector, but 3 (length of the shorter vector) does not go into 10 (length of longer vector) evenly.

```{r pg58_5}

1:3 + 1:10

```

#6

Q: What trigonometric functions does R provide?

A: Many! You can access them with ?Trig