---
title: "RFDS_CH3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(nycflights13)
library(tidyverse)
library(ggplot2)
```

When comparing two vectors for equality, use near(). "This is a safe way of comparing if two vectors of floating point numbers are (pairwise) equal. This is safer than using ==, because it has a built in tolerance"

When using filter(), & can be the same as a ,

```{r example_a}
filter(flights, arr_delay <= 120, dep_delay <= 120)
```
```{r example_a_1}
filter(flights, arr_delay <= 120 & dep_delay <= 120)
```


##### pg. 49

#1

Q: Find all flights that:

a. Had a n arrival delay of two or more hours

```{r pg49_1a}
filter(flights, arr_delay >= 120)
```

b. Flew to Houston (IAH or HOU)

```{r pg49_1b}
filter(flights, dest %in% c('IAH', 'HOU'))
```

c. Were operated by United, American or Delta

```{r pg49_1c}
filter(flights, carrier %in% c('UA', 'AA', 'DL'))
```

d. Departed in summer (July, August, and September)

```{r pg49_1d}
filter(flights, month %in% c('7', '8', '9'))
```

e. Arrived more than two hours late, but didn’t leave late

```{r pg49_1e}
filter(flights, arr_delay > 120 & dep_delay <= 0)
```

f. Were delayed by at least an hour, but made up over 30 minutes in flight

```{r pg49_1f}
filter(flights, dep_delay >= 60 & dep_delay - arr_delay > 30)
```

g. Departed between midnight and 6am (inclusive)

check min and max of dep_time

```{r pg49_1g}
filter(flights, dep_time <= 600 | dep_time == 2400)
```

#2

Q: Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?

A: ?between = "This is a shortcut for x >= left & x <= right, implemented efficiently in C++ for local values, and translated to the appropriate SQL for remote tables.
Usage
between(x, left, right)"
It would work for 1d

#3

Q: How many flights have a missing dep_time? What other variables are missing? What might these rows represent?

```{r pg49_3}
filter(flights, is.na(dep_time))
```

A: dep_delay, arr_time, arr_delay are all missing. They may be cancelled flights.

#4

Q: Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)

A: Follows mathematical rules, so the value does not matter in the examples given. For example - anything raised to 0 is going to be 1.
