---
title: "RFDS_CH1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include=FALSE, echo = FALSE}
library(ggplot2)
```

### CH 1 exercises

##### pg. 12

#1

```{r pg12_1, echo = FALSE}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy), color = "blue"
  )
```

#2

```{r pg12_2, echo = FALSE}
head(mpg)
```

#3

```{r pg12_3, echo = FALSE}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, color = displ)
  )
```

#4

```{r pg12_4, echo = FALSE}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, color = displ, size = displ)
  )
```

#5

```{r pg12_5, echo = FALSE}
?geom_point
vignette("ggplot2-specs")
```

stroke = define thickness of point border

#6

```{r pg12_6, echo = FALSE}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, color = displ < 5)
  )
```

gives a true/false!

##### pg. 15 - Facets

#1

```{r pg15_1, echo = FALSE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = drv, y = cyl)) +
  facet_grid(displ ~ cyl)
```

the plots overlay each other

#2

```{r pg15_2, echo = FALSE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = drv, y = cyl)) +
  facet_grid(drv ~ cyl)
```

the values don't exist?

#3

```{r pg15_3a, echo = FALSE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)
```

three rows by 4-wheel, front-wheel, and rear-wheel

```{r pg15_3b, echo = FALSE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```

four columns by cylinders

the period causes the plot not to facet in that direction

#4

```{r pg15_4, echo = FALSE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~ class, nrow = 2)
```

Faceting, instead of using color to distinguish data, could be helpful if you are really wanting to highlight each group as opposed to compare data between groups. If you are wanting to compare groups faceting could make it difficult for people to see the connections between graphs. A larger dataset could make one graph harder to look at because there are a lot of points

#5

nrow, ncol	- Number of rows and columns.
scales and shrink change the scales of data labels
why doesn't facet_grid have nrow/ncol? The number of rows and columns are based off of the variable used to split the data.

#6

Question: When using facet_grid, you should put the variable with the more unique levels in the columns. Why?
Answer: When putting the more levels on the row axis, then the y-axis would shrink so that it is harder to see which actual values are at the points as shown in the plot.

##### pg. 20 - geoms

#1

Question: What would you use to draw a...
line chart? geom_line
boxplot? geom_boxplot
histogram? geom_histogram
area chart? geom_area

#2

```{r pg20_2, echo = FALSE}
ggplot(data = mpg,
       mapping = aes(x = displ, y = hwy, color = drv)
       ) +
  geom_point() +
  geom_smooth(se = FALSE)
```

Colors based on drive of car - three scatter plots with lines over.

#3

show.legend = FALSE prevents the legend from showing. legend will show automatically if any aesthetics are mapped.

#4

se argument displays the confidence interval around smooth

#5

```{r pg20_5}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()
```

```{r pg20_5b}
ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

they will look the same

#6

```{r pg20_6a}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

```{r pg20_6b}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth(mapping = aes(group = drv), se = FALSE)
```

```{r pg20_6c}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = drv)) + 
  geom_smooth(se = FALSE)
```

```{r pg20_6d}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = drv)) + 
  geom_smooth(mapping = aes(linetype = drv), se = FALSE)
```

```{r pg20_6e}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = drv))
```

##### pg.26

#1

Q: What is the default geom associated with stat_summary()?
A: geom_pointrange()

Q: How can you rewrite the following using geom_pointrange()?

```{r pg26_1a}
ggplot(data = diamonds) + 
  stat_summary(mapping = aes(x = cut, y = depth),
               fun.ymin = min,
               fun.ymax = max,
               fun.y = median)
```

```{r pg26_1b}
ggplot(data = diamonds) + 
  geom_pointrange(mapping = aes(x = cut, y = depth),
                  stat = "summary",
                  fun.ymin = min,
                  fun.ymax = max,
                  fun.y = median)
```

#2

There are two types of bar charts: geom_bar() and geom_col(). geom_bar() makes the height of the bar proportional to the number of cases in each group (or if the weight aesthetic is supplied, the sum of the weights). If you want the heights of the bars to represent values in the data, use geom_col() instead. geom_bar() uses stat_count() by default: it counts the number of cases at each x position. geom_col() uses stat_identity(): it leaves the data as is.

#3

geom and stat pairs: https://ggplot2.tidyverse.org/reference/

#4

stat_smooth() calculates four variables:
  y - predicted value
  ymin - lower pointwise confidence interval around the mean
  ymax - upper pointwise confidence interval around the mean
  se - standard error
See ?stat_smooth for more details on the specific parameters. Most importantly, method controls the smoothing method to be employed, se determines whether confidence interval should be plotted, and level determines the level of confidence interval to use.

#5

If we fail to set group = 1, the proportions for each cut are calculated using the complete dataset, rather than each subset of cut.

##### pg. 31

#1

There is overplotting because there are multiple observations for certain combinations. You could use the 'jitter' position adjustment that adds random noise to each observation, therefore shifting the observations just enough to see all of the points.

```{r pg31_1}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point(position = "jitter")
```

#2

width and height control the amount of jittering. The jitter is added in both positive and negative directions, so the total spread is twice the value specified here.

#3

geom_count adjusts the size of the point to show how many observations are at that location. geom_jitter adds random noise to each point to slightly shift them and make each point observable.

#4

Q: What's the default position adjustment for geom_boxplot()?
A: position = "dodge2".
Looking at ?position_dodge2 - "Dodging preserves the vertical position of an geom while adjusting the horizontal position. position_dodge2 is a special case of position_dodge for arranging box plots, which can have variable widths. position_dodge2 also works with bars and rectangles."

```{r pg31_4}
ggplot(data = mpg, aes(x = displ, y = hwy, color = drv)) +
  geom_boxplot()
```

##### pg. 33

#1

Q: Turn a stacked bar chart into a pie chart using coord_polar()

```{r pg33_1}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = factor(1), fill = clarity))
```

```{r pg33_1b}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = factor(1), fill = clarity)) +
  coord_polar(theta = "y")
```

from https://jrnold.github.io/r4ds-exercise-solutions/data-visualisation.html#exercise-3.8.4: The argument theta = "y" maps y to the angle of each section. If coord_polar() is specified without theta = "y", then the resulting plot is called a bulls-eye chart.

#2

Q: What does labs() do?
A: Controls the labels on your plots. ?labs()... "Good labels are critical for making your plots accessible to a wider audience. Always ensure the axis and legend labels display the full variable name. Use the plot title and subtitle to explain the main findings. It's common to use the caption to provide information about the data source. tag can be used for adding identification tags to differentiate between multiple plots."

#3

Q: What is the difference between coord_quickmap() and coord_map()?
A: ?coord_map... "coord_map projects a portion of the earth, which is approximately spherical, onto a flat 2D plane using any projection defined by the mapproj package. Map projections do not, in general, preserve straight lines, so this requires considerable computation. coord_quickmap is a quick approximation that does preserve straight lines. It works best for smaller areas closer to the equator."

#4

Q: What does the following plot tell you about the relationship between city and highway mpg? Why is coord_fixed() important? What does geom_abline() do?

```{r pg33_4}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() +
  geom_abline() +
  coord_fixed()
```

A: hwy mpg rise faster than city mpg. The coord_fixed() is important because it sets the coordinates as equal, so you are actually comparing two like things. geom_abline() adds a reference line that is at 45 degrees (due to the coord_fixed()).